### 2022-09-14 Manuscript Code.Rmd
Edited for EHP peer review

---
title: "Manuscript Code"
author: "Kaya Bramble"
date: "5/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
    res <- suppressWarnings(
        lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
               detach, character.only=TRUE, unload=TRUE, force=TRUE))

}

pacman::p_load(dplyr, ggplot2, ggpubr, sf, tidyverse, modelr, nlme, gstat, sp, geepack, kableExtra)

```

## Preparing Data

```{r}

## Air pollution predictions at block centroids
block_preds <- readRDS("/projects/trap/transfer/kaya/Code Review/Block Predictions/20220207/predictions.rda")

## 2010 census data at block level
census_block <- read.csv("/projects/trap/transfer/kaya/Code Review/dr0342_census_block.csv")

## 2010 census data at block group level
census_blockgroup <- read.csv("/projects/trap/transfer/kaya/Code Review/dr0342_census_blockgroup.csv")

## Corresponding HOLC grades to blocks
holc_grades <- read.csv("/projects/trap/transfer/kaya/Code Review/holc_grades.csv")

## Blocks wihin redlining area
redlining_area <- read.csv("/projects/trap/transfer/kaya/Code Review/redlining_area.csv")

## Convert native id column to block column
block_preds$block_key <- substr(block_preds$native_id, 1, 15)
block_preds$blockgroup_key <- substr(block_preds$native_id, 1, 12)

## Remove blocks not in monitoring area
block_preds <- subset(block_preds, in_monitoring_area == TRUE)

## Select data used
block_preds <- subset(block_preds, variable == "ma200_ir_bc1" | variable == "pm2.5_ug_m3" | variable == "no2" | variable == "pnc_noscreen")
block_preds <- block_preds[c("block_key", "blockgroup_key", "variable", "prediction", "longitude", "latitude")]

census_block <- census_block[c("block_key", "bk_tot_pop_2010", "bk_p_race_black", "bk_p_race_asian", "bk_p_race_native", "bk_p_race_pacific", "bk_p_ethn_non_hisp_white", "bk_p_ethn_hisp", "households")]

census_blockgroup <- census_blockgroup[c("blockgroup_key", "bg_med_inc_hshld")]

## Merge census data with predictions
census_block$block_key <- as.character(census_block$block_key)
census_blockgroup$blockgroup_key <- as.character(census_blockgroup$blockgroup_key)
census_preds <- block_preds %>%
  left_join(census_block, by="block_key") %>%
  left_join(census_blockgroup, by="blockgroup_key")  %>%
  filter(households != 0) %>%
  filter(bk_tot_pop_2010 != 0) %>%
  filter(!is.na(bg_med_inc_hshld))

## Define income groups
census_preds$income_group <- cut(census_preds$bg_med_inc_hshld, c(0, 20000, 40000, 50000, 60000, 70000, 80000, 90000, 110000, 250001))
levels(census_preds$income_group) <- c("<$20k", "$20k-40k", "$40k-50k", "$50k-60k", "$60k-70k", "$70k-80k", "$80k-90k", "$90k-110k", ">$110k")

## Create wide dataframe
census_preds_wide <- census_preds %>%
  pivot_wider(names_from=variable, values_from=prediction)

## Save file
write.csv(census_preds,"/projects/trap/transfer/kaya/census_preds.csv", row.names = FALSE)

## Assign A-D HOLC grades
holc_grades <- distinct(holc_grades)
holc_grades$BLOCKID10 <- as.character(holc_grades$BLOCKID10)
redlining_blocks <- census_preds_wide %>%
  left_join(holc_grades, by=c("block_key"="BLOCKID10"))
redlining_blocks$count <- as.numeric(ave(redlining_blocks$block_key, redlining_blocks$block_key, FUN = length))

## Assign X HOLC grade
redlining_blocks$holc_grade[is.na(redlining_blocks$holc_grade) & redlining_blocks$block_key %in% redlining_area$BLOCKID10] <- "X"
redlining_blocks <- na.omit(redlining_blocks)

## Save file
write.csv(redlining_blocks,"/projects/trap/transfer/kaya/redlining_blocks.csv", row.names = FALSE)

## Percent Different Function
per_diff <- function(x1, x2) {
  diff <- (x1 - x2)/((x1 + x2)/2) * 100
  diff <- round(diff, digits=0)
  return(diff)
}

```

## Percentage of without households/household income data

```{r}

census_percent <- block_preds %>%
  left_join(census_block, by="block_key") %>%
  left_join(census_blockgroup, by="blockgroup_key") %>%
  filter(bk_tot_pop_2010 != 0)

census_percent_wide <- census_percent %>%
  pivot_wider(names_from=variable, values_from=prediction)

nrow(subset(census_percent_wide, households == 0))

nrow(subset(census_percent_wide, households == 0)) / nrow(census_percent_wide) * 100

sum(subset(census_percent_wide, households == 0)$bk_tot_pop_2010) / sum(census_percent_wide$bk_tot_pop_2010) * 100

nrow(subset(census_percent_wide, is.na(bg_med_inc_hshld) & households!= 0))

nrow(subset(census_percent_wide, is.na(bg_med_inc_hshld) & households!= 0)) / nrow(census_percent_wide) * 100

sum(subset(census_percent_wide, is.na(bg_med_inc_hshld) & households!= 0)$bk_tot_pop_2010) / sum(census_percent_wide$bk_tot_pop_2010) * 100

```

## Table 1. Summary Statistics

```{r}

outside_redlining <- subset(census_preds_wide, !(census_preds_wide$block_key %in% redlining_blocks$block_key))

## n Blocks
nrow(census_preds_wide)
length(unique(redlining_blocks$block_key))
nrow(census_preds_wide) - length(unique(redlining_blocks$block_key))
redlining_blocks %>%
  group_by(holc_grade) %>%
  summarise(pop = sum(1/count))

## n People
sum(census_preds_wide$bk_tot_pop_2010)
sum(redlining_blocks$bk_tot_pop_2010/redlining_blocks$count)
sum(census_preds_wide$bk_tot_pop_2010) - sum(redlining_blocks$bk_tot_pop_2010/redlining_blocks$count)
redlining_blocks %>%
  group_by(holc_grade) %>%
  summarise(pop = sum(bk_tot_pop_2010/count))

## Pollutants and demographics
summary(census_preds_wide)
summary(data.frame(unique(redlining_blocks)))
summary(outside_redlining)
summary(subset(redlining_blocks, holc_grade == "A"))
summary(subset(redlining_blocks, holc_grade == "B"))
summary(subset(redlining_blocks, holc_grade == "C"))
summary(subset(redlining_blocks, holc_grade == "D"))
summary(subset(redlining_blocks, holc_grade == "X"))

```

#### Number of blocks split between redlining grades

```{r}

## Table of blocks and how many times they are graded
n_occur <- data.frame(table(redlining_blocks$block_key))

## Total number of blocks in HOLC-graded area
nrow(n_occur)

## Total number of blocks split 
nrow(n_occur[n_occur$Freq > 1,])

## Number of blocks split between 3 grades
nrow(n_occur[n_occur$Freq > 2,])

```

## Table 2. Exposure Disparities by Income

```{r, fig.show="hold", out.width="50%", echo=FALSE}

census_preds_wide %>%
  group_by(income_group) %>%
  summarise(pop = sum(bk_tot_pop_2010),
            ufp = weighted.mean(pnc_noscreen, bk_tot_pop_2010),
            ufp_pct_diff = per_diff(ufp, 5712),
            bc = weighted.mean(ma200_ir_bc1, bk_tot_pop_2010),
            bc_pct_diff = per_diff(bc, 453),
            no2 = weighted.mean(no2, bk_tot_pop_2010),
            no2_pct_diff = per_diff(no2, 7.63),
            pm2.5 = weighted.mean(pm2.5_ug_m3, bk_tot_pop_2010),
            pm2.5_pct_diff = per_diff(pm2.5, 4.27))

```

## Table 3. Exposure Disparities by Race-Ethnicity

```{r}

census_preds_long <- census_preds_wide %>%
  pivot_longer(cols = c(bk_p_race_asian, bk_p_race_black, bk_p_ethn_hisp, bk_p_ethn_non_hisp_white, bk_p_race_native, bk_p_race_pacific), names_to="race", values_to="percent")

weighted.mean(census_preds_wide$pnc_noscreen, census_preds_wide$bk_tot_pop_2010)
weighted.mean(census_preds_wide$ma200_ir_bc1, census_preds_wide$bk_tot_pop_2010)
weighted.mean(census_preds_wide$no2, census_preds_wide$bk_tot_pop_2010)
weighted.mean(census_preds_wide$pm2.5_ug_m3, census_preds_wide$bk_tot_pop_2010)

census_preds_long %>%
  group_by(race) %>%
  summarise(pop = sum(percent * bk_tot_pop_2010 / 100),
            ufp = weighted.mean(pnc_noscreen, bk_tot_pop_2010 * percent),
            ufp_pct_diff = per_diff(ufp, 6678),
            bc = weighted.mean(ma200_ir_bc1, bk_tot_pop_2010 * percent),
            bc_pct_diff = per_diff(bc, 520),
            no2 = weighted.mean(no2, bk_tot_pop_2010 * percent),
            no2_pct_diff = per_diff(no2, 8.74),
            pm2.5 = weighted.mean(pm2.5_ug_m3, bk_tot_pop_2010 * percent),
            pm2.5_pct_diff = per_diff(pm2.5, 4.56))

```

## Table 4. Exposure Disparities by Redlining Grade

```{r}

## Pollution means weighted by block population divided by number of grades block is assigned to
redlining_blocks %>%
  group_by(holc_grade) %>%
  summarise(ufp = weighted.mean(pnc_noscreen, bk_tot_pop_2010/count),
            ufp_pct_diff = per_diff(ufp, 7150),
            bc = weighted.mean(ma200_ir_bc1, bk_tot_pop_2010/count),
            bc_pct_diff = per_diff(bc, 547),
            no2 = weighted.mean(no2, bk_tot_pop_2010/count),
            no2_pct_diff = per_diff(no2, 9.29),
            pm2.5 = weighted.mean(pm2.5_ug_m3, bk_tot_pop_2010/count),
            pm2.5_pct_diff = per_diff(pm2.5, 4.75))

```

## GEE Models

```{r}

# Created nested dfs by pollutant
by_pollutant <- census_preds %>% 
  group_by(variable) %>% 
  nest()

# Create model function
pollutant_model <- function(df) {
  lm(prediction ~ log10(bg_med_inc_hshld), weights = bk_tot_pop_2010, data=df)
}

## Apply function to each group (pollutant)
by_pollutant <- by_pollutant %>% 
  mutate(model = map(data, pollutant_model),
         resids = map2(data, model, add_residuals))

## Unnest with broom::glance  
by_pollutant %>% 
  mutate(glance=map(model, broom::glance)) %>% 
  unnest(glance)

by_pollutant %>% 
  mutate(tidy = map(model, broom::tidy)) %>%
  unnest(tidy) %>% 
  dplyr::select(term, estimate, std.error) %>% 
  mutate(estimate = estimate*log10(1.1))

# Create dummy grouping variable  
census_preds$group = 1

census_preds_pnc = census_preds %>% 
   filter(variable == "pnc_noscreen")

model_nosp_pnc <- lme(fixed = prediction ~ log10(bg_med_inc_hshld), 
                  data = census_preds_pnc, random = ~ 1|group, method = "ML",
                  weights = ~bk_tot_pop_2010)

summary(model_nosp_pnc)

census_preds_dist = census_preds %>% 
  select(block_key, bk_tot_pop_2010, bg_med_inc_hshld, longitude, latitude) %>% 
  distinct()

df_resids_locs = census_preds_pnc %>% 
  select(block_key) %>% 
  mutate(block_key = as.character(block_key)) %>% 
  left_join(census_preds_dist, by=c("block_key" = "block_key"))

resids = model_nosp_pnc$residuals[,1]

df_resids_locs$resids_pnc = resids

coordinates(df_resids_locs) = ~longitude+latitude
# vario = variogram(resids_pnc~longitude+latitude, df_resids_locs)
# plot(vario)

rm(df_resids_locs)
sites <- sample(1:16386, 200)
sampledata <- census_preds_pnc[sites, ]

null_model <- lme(fixed = prediction ~ log10(bg_med_inc_hshld), 
                  data = sampledata, random = ~ 1|group)
summary(null_model)

exp.sp <- update(null_model, correlation = corExp(1, form= ~ longitude+latitude))
summary(exp.sp)

mod_ind_full <- geeglm(formula = prediction ~ log10(bg_med_inc_hshld),
                  data = census_preds_pnc, 
                  weights = bk_tot_pop_2010,
                  id=block_key, 
                  corstr="independence")

summary(mod_ind_full)

tidy(mod_ind_full, conf.int = TRUE)

# Created nested dfs by pollutant
mhi_pollutant <- census_preds %>% 
  group_by(variable) %>% 
  nest()

# Create model function
mhi_model <- function(df) {
  geeglm(formula = prediction ~ log10(bg_med_inc_hshld),
                  data = df, 
                  weights = bk_tot_pop_2010,
                  id=block_key, 
                  corstr="independence")
}

## Apply function to each group (pollutant)
mhi_pollutant <- mhi_pollutant %>% 
  mutate(model = map(data, mhi_model),
         resids = map2(data, model, add_residuals))

inc_plots <- mhi_pollutant %>% 
  mutate(tidy = map(model, broom::tidy)) %>%
  unnest(tidy) %>% 
  dplyr::select(term, estimate, std.error) %>% 
  filter(term=="log10(bg_med_inc_hshld)") %>% 
  select(-term) %>% 
  mutate(conf.lo = estimate - 1.96*std.error, 
         conf.hi = estimate + 1.96*std.error) %>% 
  mutate(estimate = estimate*log10(1.1),
         std.error = std.error*log10(1.1),
         conf.lo = conf.lo*log10(1.1),
         conf.hi = conf.hi*log10(1.1))

mhi_pollutant %>% 
  mutate(tidy = map(model, broom::tidy)) %>%
  unnest(tidy) %>% 
  dplyr::select(term, estimate, std.error) %>% 
  filter(term=="log10(bg_med_inc_hshld)") %>% 
  select(-term) %>% 
  mutate(conf.lo = estimate - 1.96*std.error, 
         conf.hi = estimate + 1.96*std.error) %>% 
  mutate(estimate = estimate*log10(1.1),
         std.error = std.error*log10(1.1),
         conf.lo = conf.lo*log10(1.1),
         conf.hi = conf.hi*log10(1.1)) %>% 
  kable() %>% 
  kable_styling()

# Created nested dfs by pollutant
race_pollutant <- census_preds %>% 
  pivot_longer(cols=c("bk_p_ethn_non_hisp_white", "bk_p_race_asian", "bk_p_race_black", "bk_p_ethn_hisp",
                      "bk_p_race_pacific", "bk_p_race_native"),
               names_to = "race_eth_grp",
               values_to = "bk_p_race") %>% 
  group_by(variable, race_eth_grp) %>% 
  nest()

# Create model function
perc_race_model <- function(df) {
  geeglm(formula = prediction ~ bk_p_race,
                  data = df, 
                  weights = bk_tot_pop_2010,
                  id=block_key, 
                  corstr="independence")
}

## Apply function to each group (pollutant)
perc_race <- race_pollutant %>% 
  mutate(model = map(data, perc_race_model),
         resids = map2(data, model, add_residuals))

perc_race %>% 
  mutate(tidy = map(model, broom::tidy)) %>%
  unnest(tidy) %>% 
  dplyr::select(term, estimate, std.error) %>% 
  select(-term) %>% 
  mutate(conf.lo = estimate - 1.96*std.error, 
         conf.hi = estimate + 1.96*std.error) %>% 
  kable() %>% 
  kable_styling()

fig_4 <- data.frame(perc_race %>% 
  mutate(tidy = map(model, broom::tidy)) %>%
  unnest(tidy) %>% 
  dplyr::select(term, estimate, std.error) %>% 
  filter(term =="bk_p_race") %>%
  filter(race_eth_grp !="bk_p_race_pacific") %>%
  filter(race_eth_grp !="bk_p_race_native") %>%
  select(-term) %>% 
  mutate(conf.lo = estimate - 1.96*std.error, 
         conf.hi = estimate + 1.96*std.error))

fig_4$variable <- recode(fig_4$variable, pnc_noscreen = "UFP (pt/cm3)", ma200_ir_bc1 = "BC (µg/m3)", no2 = "NO2 (ppb)", pm2.5_ug_m3 = "PM2.5 (µg/m3)", ) 
fig_4$race_eth_grp <- recode(fig_4$race_eth_grp, bk_p_ethn_non_hisp_white = "Non-Hispanic White", bk_p_race_asian = "Asian", bk_p_race_black = "Black", bk_p_ethn_hisp = "Hispanic")

# Created nested dfs by pollutant
holc_df <- redlining_blocks %>% 
  pivot_longer(cols=c("ma200_ir_bc1", "no2", "pm2.5_ug_m3", "pnc_noscreen"),
               names_to="variable",
               values_to="prediction") %>% 
  group_by(variable) %>% 
  nest()

holc_model <- function(df) {
  geeglm(formula = prediction ~ holc_grade,
                  data = df, 
                  weights = bk_tot_pop_2010 / count,
                  id=block_key, 
                  corstr="independence")
}

## Apply function to each group (pollutant)
holc_output <- holc_df %>% 
  mutate(model = map(data, holc_model),
         resids = map2(data, model, add_residuals))

holc_output %>% 
  mutate(tidy = map(model, broom::tidy)) %>%
  unnest(tidy) %>% 
  dplyr::select(term, estimate, std.error) %>% 
  mutate(conf.lo = estimate - 1.96*std.error, 
         conf.hi = estimate + 1.96*std.error) %>% 
  kable() %>%
  kable_styling()

# Centering by holc grade A gives us intercepts close to those from single predictor model for holc grade
mlr_df <- redlining_blocks %>% 
  mutate(adjusted_inc = log10(bg_med_inc_hshld) - log10(117134),
        adjusted_race = bk_p_ethn_non_hisp_white -85.82) %>%
  pivot_longer(cols=c("ma200_ir_bc1", "no2", "pm2.5_ug_m3", "pnc_noscreen"),
               names_to="variable",
               values_to="prediction") %>% 
  group_by(variable) %>% 
  nest()

mlr_model <- function(df) {
  geeglm(formula = prediction ~ adjusted_inc + adjusted_race + holc_grade,
                  data = df, 
                  weights = bk_tot_pop_2010 / count,
                  id=block_key, 
                  corstr="independence")
}

## Apply function to each group (pollutant)
mlr_output <- mlr_df %>% 
  mutate(model = map(data, mlr_model),
         resids = map2(data, model, add_residuals))

mlr_output %>% 
  mutate(tidy = map(model, broom::tidy)) %>%
  unnest(tidy) %>% 
  dplyr::select(term, estimate, std.error) %>% 
  mutate(conf.lo = estimate - 1.96*std.error, 
         conf.hi = estimate + 1.96*std.error) %>% 
  mutate(estimate = ifelse(term == "adjusted_inc", estimate*log10(1.1), estimate),
          std.error = ifelse(term == "adjusted_inc", std.error*log10(1.1), std.error),
          conf.lo = ifelse(term == "adjusted_inc", conf.lo*log10(1.1), conf.lo),
          conf.hi = ifelse(term == "adjusted_inc", conf.hi*log10(1.1), conf.hi)) %>%
  mutate(estimate = ifelse(term == "adjusted_race", estimate*10, estimate),
          std.error = ifelse(term == "adjusted_race", std.error*10, std.error),
          conf.lo = ifelse(term == "adjusted_race", conf.lo*10, conf.lo),
          conf.hi = ifelse(term == "adjusted_race", conf.hi*10, conf.hi)) %>%
  kable() %>% 
  kable_styling()

fig_4$race_eth_grp <- factor(fig_4$race_eth_grp, levels = c("Non-Hispanic White", "Asian", "Hispanic", "Black"))
fig_4 <- fig_4 %>%
  group_by(variable) %>%
  mutate(low = -(max(abs(conf.hi*5))), high = max(conf.hi*5))

ggplot(data=fig_4, aes(x=race_eth_grp, y=estimate*5, ymin=conf.lo*5, ymax=conf.hi*5)) +
  geom_pointrange(aes(col=race_eth_grp)) + 
  geom_hline(aes(col=race_eth_grp), yintercept=0, linetype=2) +
  xlab("Pollutant") + ylab("Mean Difference (95% CI) per 5% Increase in Population") + 
  geom_errorbar(aes(ymin=conf.lo*5, ymax=conf.hi*5, col=race_eth_grp), width=0.5 ,cex=1) + 
  facet_wrap(~factor(variable, levels=c("UFP (pt/cm3)","BC (ng/m3)","NO2 (ppb)","PM2.5 (µg/m3)")), strip.position="left", nrow=4, scales = "free_x") +
  geom_blank(aes(y = low)) +
  geom_blank(aes(y = high)) +
  theme(plot.title=element_text(size=16,face="bold"),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      axis.text.x=element_text(face="bold"),
      axis.title=element_text(size=12,face="bold"),
      strip.text.y = element_text(hjust=0,vjust = 1,angle=180,face="bold")) +
  labs(color="Race-Ethnicity") +
  coord_flip()

```

```{r, fig.show="hold", out.width="50%"}

ggplot(census_preds_wide, aes(x=log10(bg_med_inc_hshld), y=ma200_ir_bc1)) + geom_point() + geom_abline(slope=(-12.1)/log10(1.1), intercept=79.9/log10(1.1), color="red") + geom_smooth()

ggplot(census_preds_wide, aes(x=log10(bg_med_inc_hshld), y=no2)) + geom_point() + geom_abline(slope=(-.224)/log10(1.1), intercept=1.442/log10(1.1), color="red") + geom_smooth()

ggplot(census_preds_wide, aes(x=log10(bg_med_inc_hshld), y=pm2.5_ug_m3)) + geom_point() + geom_abline(slope=(-.0407)/log10(1.1), intercept=0.385/log10(1.1), color="red") + geom_smooth()

ggplot(census_preds_wide, aes(x=log10(bg_med_inc_hshld), y=pnc_noscreen)) + geom_point() + geom_abline(slope=(-198.04)/log10(1.1), intercept=1232.50/log10(1.1), color="red") + geom_smooth()

ggplot(census_preds_wide, aes(x=bk_p_race_asian, y=pnc_noscreen)) + geom_point() + geom_smooth() + geom_abline(slope=11.494, intercept=6546.948, color="red")
ggplot(census_preds_wide, aes(x=bk_p_race_black, y=pnc_noscreen)) + geom_point() + geom_smooth() + geom_abline(slope=77.238, intercept=6229.281, color="red")
ggplot(census_preds_wide, aes(x=bk_p_ethn_hisp, y=pnc_noscreen)) + geom_point() + geom_smooth() + geom_abline(slope=34.451, intercept=6398.682, color="red")
ggplot(census_preds_wide, aes(x=bk_p_ethn_non_hisp_white, y=pnc_noscreen)) + geom_point() + geom_smooth() + geom_abline(slope=-30.901, intercept=8694.838, color="red")

ggplot(census_preds_wide, aes(x=bk_p_race_asian, y=ma200_ir_bc1)) + geom_point() + geom_smooth() + geom_abline(slope=-0.038, intercept=522.568, color="red")
ggplot(census_preds_wide, aes(x=bk_p_race_black, y=ma200_ir_bc1)) + geom_point() + geom_smooth() + geom_abline(slope=3.359, intercept=500.701, color="red")
ggplot(census_preds_wide, aes(x=bk_p_ethn_hisp, y=ma200_ir_bc1)) + geom_point() + geom_smooth() + geom_abline(slope=1.204, intercept=510.809, color="red")
ggplot(census_preds_wide, aes(x=bk_p_ethn_non_hisp_white, y=ma200_ir_bc1)) + geom_point() + geom_smooth() + geom_abline(slope=-1.053, intercept=589.312, color="red")

ggplot(census_preds_wide, aes(x=bk_p_race_asian, y=no2)) + geom_point() + geom_smooth() + geom_abline(slope=-0.002, intercept=8.803, color="red")
ggplot(census_preds_wide, aes(x=bk_p_race_black, y=no2)) + geom_point() + geom_smooth() + geom_abline(slope=0.048, intercept=8.466, color="red")
ggplot(census_preds_wide, aes(x=bk_p_ethn_hisp, y=no2)) + geom_point() + geom_smooth() + geom_abline(slope=0.007, intercept=8.705, color="red")
ggplot(census_preds_wide, aes(x=bk_p_ethn_non_hisp_white, y=no2)) + geom_point() + geom_smooth() + geom_abline(slope=-0.012, intercept=9.509, color="red")

ggplot(census_preds_wide, aes(x=bk_p_race_asian, y=pm2.5_ug_m3)) + geom_point() + geom_smooth() + geom_smooth() + geom_abline(slope=-0.004, intercept=4.616, color="red")
ggplot(census_preds_wide, aes(x=bk_p_race_black, y=pm2.5_ug_m3)) + geom_point() + geom_smooth() + geom_smooth() + geom_abline(slope=0.010, intercept=4.496, color="red")
ggplot(census_preds_wide, aes(x=bk_p_ethn_hisp, y=pm2.5_ug_m3)) + geom_point() + geom_smooth() + geom_smooth() + geom_abline(slope=0.001, intercept=4.551, color="red")
ggplot(census_preds_wide, aes(x=bk_p_ethn_non_hisp_white, y=pm2.5_ug_m3)) + geom_point() + geom_smooth() + geom_smooth() + geom_abline(slope=-0.001, intercept=4.642, color="red")

```

# Supplement ##################################################################

### Sensitivity Analysis: HOLC-Graded Area Definitions

```{r}

## Assign A-D HOLC grades - Analysis 1
holc_grades$duplicated <- duplicated(holc_grades$BLOCKID10)
between_grades <- holc_grades[duplicated(holc_grades$BLOCKID10), ]$BLOCKID10
holc_no_dup <- holc_grades[!(holc_grades$BLOCKID10 %in% between_grades), ]

holc_no_dup$BLOCKID10 <- as.character(holc_no_dup$BLOCKID10)
redlining_blocks <- census_preds_wide %>%
  left_join(holc_no_dup, by=c("block_key"="BLOCKID10"))
redlining_blocks$count <- as.numeric(ave(redlining_blocks$block_key, redlining_blocks$block_key, FUN = length))

## Assign X HOLC grade - Analysis 1
redlining_blocks$holc_grade[is.na(redlining_blocks$holc_grade) & redlining_blocks$block_key %in% redlining_area$BLOCKID10] <- "X"
redlining_blocks <- na.omit(redlining_blocks)

## Assign A-D HOLC grades - Analysis 2
holc_areas <- read.csv("/projects/trap/transfer/kaya/area definition.csv")
holc_areas <- holc_areas[c("BLOCKID10", "holc_grade", "area")]
holc_area_per <- holc_areas %>%
    group_by(BLOCKID10) %>%
    mutate(percent = area/sum(area))

holc_area_per$BLOCKID10 <- as.character(holc_area_per$BLOCKID10)
redlining_blocks_area <- census_preds_wide %>%
  left_join(holc_area_per, by=c("block_key"="BLOCKID10"))

## Assign X HOLC grade - Analysis 2
redlining_blocks_area$holc_grade[is.na(redlining_blocks_area$holc_grade) & redlining_blocks_area$block_key %in% redlining_area$BLOCKID10] <- "X"
redlining_blocks_area <- na.omit(redlining_blocks_area)

## Analysis 1: Pollution means weighted by block population divided by number of grades block is assigned to
redlining_blocks %>%
  group_by(holc_grade) %>%
  summarise(ufp = weighted.mean(pnc_noscreen, bk_tot_pop_2010/count),
            ufp_pct_diff = per_diff(ufp, 7102),
            bc = weighted.mean(ma200_ir_bc1, bk_tot_pop_2010/count),
            bc_pct_diff = per_diff(bc, 539),
            no2 = weighted.mean(no2, bk_tot_pop_2010/count),
            no2_pct_diff = per_diff(no2, 9.3),
            pm2.5 = weighted.mean(pm2.5_ug_m3, bk_tot_pop_2010/count),
            pm2.5_pct_diff = per_diff(pm2.5, 4.7))

## Analysis 2: Pollution means weighted by block population divided by number of grades block is assigned to
redlining_blocks_area %>%
  group_by(holc_grade) %>%
  summarise(ufp = weighted.mean(pnc_noscreen, bk_tot_pop_2010*percent),
            ufp_pct_diff = per_diff(ufp, 7102),
            bc = weighted.mean(ma200_ir_bc1, bk_tot_pop_2010*percent),
            bc_pct_diff = per_diff(bc, 539),
            no2 = weighted.mean(no2, bk_tot_pop_2010*percent),
            no2_pct_diff = per_diff(no2, 9.3),
            pm2.5 = weighted.mean(pm2.5_ug_m3, bk_tot_pop_2010*percent),
            pm2.5_pct_diff = per_diff(pm2.5, 4.7))

```

### Sensitivity Analysis: Results Restricted to the HOLC-Graded Area

``` {r}

redlining_blocks_unique <- subset(redlining_blocks, !duplicated(redlining_blocks$block_key))

redlining_blocks_long <- redlining_blocks_unique %>%
  pivot_longer(cols = c(bk_p_race_asian, bk_p_race_black, bk_p_ethn_hisp, bk_p_ethn_non_hisp_white, bk_p_race_native, bk_p_race_pacific), names_to="race", values_to="percent")

redlining_blocks_unique %>%
  group_by(income_group) %>%
  summarise(pop = sum(bk_tot_pop_2010),
            ufp = weighted.mean(pnc_noscreen, bk_tot_pop_2010),
            ufp_pct_diff = per_diff(ufp, 7192),
            bc = weighted.mean(ma200_ir_bc1, bk_tot_pop_2010),
            bc_pct_diff = per_diff(bc, 557),
            no2 = weighted.mean(no2, bk_tot_pop_2010),
            no2_pct_diff = per_diff(no2, 9.76),
            pm2.5 = weighted.mean(pm2.5_ug_m3, bk_tot_pop_2010),
            pm2.5_pct_diff = per_diff(pm2.5, 4.92))

weighted.mean(redlining_blocks_unique$pnc_noscreen, redlining_blocks_unique$bk_tot_pop_2010)
weighted.mean(redlining_blocks_unique$ma200_ir_bc1, redlining_blocks_unique$bk_tot_pop_2010)
weighted.mean(redlining_blocks_unique$no2, redlining_blocks_unique$bk_tot_pop_2010)
weighted.mean(redlining_blocks_unique$pm2.5_ug_m3, redlining_blocks_unique$bk_tot_pop_2010)

redlining_blocks_long %>%
  group_by(race) %>%
  summarise(pop = sum(percent * bk_tot_pop_2010 / 100),
            ufp = weighted.mean(pnc_noscreen, bk_tot_pop_2010 * percent),
            ufp_pct_diff = per_diff(ufp, 8232),
            bc = weighted.mean(ma200_ir_bc1, bk_tot_pop_2010 * percent),
            bc_pct_diff = per_diff(bc, 625),
            no2 = weighted.mean(no2, bk_tot_pop_2010 * percent),
            no2_pct_diff = per_diff(no2, 11.4),
            pm2.5 = weighted.mean(pm2.5_ug_m3, bk_tot_pop_2010 * percent),
            pm2.5_pct_diff = per_diff(pm2.5, 4.99))

```